---
title: "answers02.Rmd"
author: "Matt Doherty"
date: "April 17, 2015"
output: html_document
---
```{r}
metadata <- read.table(file="~/Assignment02 git/wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]

n.total <- nrow(metadata) #total rows == total samples

```
1.  Calculate the following on the data we read in from the `wild.metadata.txt` file that we discussed in class on 9/16/2014.

  * How many samples were described in the `wild.metadata.txt`? 
  
  There were *__`r n.total`__* samples described.
  
  * How many columns are in the table? What are their names?
  
  There are *__`r ncol(metadata)`__* columns in the table.
  Their names are *__`r colnames(metadata)`__*
  

  
  * How many samples came from mice that weighed 15 or more grams?
```{r}
five.teen <- metadata$Weight >= "15.0" #all samples weighing 15.0g or more
n.grfiveteen <- nrow(metadata[five.teen,]) #number of samples 15.0g or more
```
  There were *__`r n.grfiveteen`__* samples from mice >= 15.0g.
  
  
```{r}
ave.weight <- sum(metadata$Weight)/n.total #average weight
med.weight <- median(metadata$Weight)  #median
```
  * What is the median weight of the samples?
  *__`r med.weight`__* is the median weight of all samples and *__`r ave.weight`__* is the average.
  
```{r}
tot.PMG <- metadata$SP=="PMG" #all PMG mice
n.PMG <- nrow(metadata[tot.PMG,]) #number of PNG mice

Fem.PL <- metadata$SP == "PL" & metadata$Sex == "F"  #female PL mice 
n.fempl <- nrow(metadata[Fem.PL,]) #number of female PL mice
```  
  * How many PMG samples were there?
  
  There were *__`r n.PMG`__* PMG mice samples.
  
  * How many female PL samples were there?
  
  There were *__`r n.fempl`__* female PL samples.
 
```{r}
order.ET <- order(metadata$ET) #order of samples by ET
sortby.ET <- metadata[order.ET, ] #metadata sorted by ET
sortby.ET15 <- sortby.ET[1:5, ] #shows rows 1-5 sorted by ET
```
  * Alphabetize `wild.metadata.txt` by the ear tag number (only show the first 5 rows of the table)
```{r}
sortby.ET15
```

  
  * Sort the table by the weight of the mice that each sample came from
```{r}
order.Weight <- order(metadata$Weight) #order of samples by Weight
sortby.Weight <- metadata[order.Weight, ] #metadata sorted by Weight
sortby.Weight[1:5, ]
```

  * The `Station` column indicates where the mice were sampled. Where were the most mice captured?
```{r}
order.Station <- order(metadata$Station) #order of samples by ET
sortby.Station <- metadata[order.Station, ]
summary(sortby.Station)
mode.Station <- "N20"
tot.N20 <- metadata$Station=="N20" #all Station N20 mice
n.N20 <- nrow(metadata[tot.N20, ]) #number of Station N20 mice

modeav <- function (x, method = "mode", na.rm = FALSE)
{
    x <- unlist(x)
    if (na.rm)
    x <- x[!is.na(x)]
    u <- unique(x)
    n <- length(u)
#get frequencies of each of the unique values in the vector
    frequencies <- rep(0, n)
    for (i in seq_len(n)) {
        if (is.na(u[i])) {
            frequencies[i] <- sum(is.na(x))
        }
        else {
            frequencies[i] <- sum(x == u[i], na.rm = TRUE)
        }
    }
#mode if a unimodal vector, else NA
    if (method == "mode" | is.na(method) | method == "")
        {return(ifelse(length(frequencies[frequencies==max(frequencies)])>1,NA,u[which.max(frequencies)]))}
#number of modes
    if (method == "modes" | method == "modevalues")
        {return(length(frequencies[frequencies==max(frequencies)]))}
#list of all modes
    if(method == "nmode" | method == "nmodes")
        {return(u[which(frequencies==max(frequencies), arr.ind = FALSE, useNames = FALSE)])}
#error trap the method
    warning("Warning: method not recognised.  Valid methods are 'mode' [default], 'nmodes' and 'modes'")
    return()
} #found at 'http://stackoverflow.com/revisions/15620958/1'

mode.Station <- modeav(metadata$Station, method="nmode") #reassigns mode.Station using modeav function
```
The most mice were captured at *__Station `r mode.Station`__*

  * How many mice were captured there?

There were *__`r n.N20`__* captured there.

2.	Describe what each of the following commands does in a text block above the code, be specific. Put the code into the appropriate knitr code chunk. If something throws an error or looks weird, fix it.

Creates a sequence of numbers from 1 to 100 by 3s.
```{r}
seq(1,100,3)
```

Creates 10 repeats of "a, b"
```{r}
rep(c("a","b"),10)
```

Creates a random list of numbers from 1-10
```{r}
r <- runif(10); order(r)
```

Gives the remainder of dividing 100 by 3
```{r}
100 %% 3
```

Displays a table from 'wild.metadata.txt' featuring all samples from PMG mice weighing 16g

```{r}
metadata[metadata$Weight==16 & metadata$SP=="PMG",]
```


3.	Calculate the mode for the weights in `wild.metadata.txt`
```{r} 
modeav <- function (x, method = "mode", na.rm = FALSE)
{
    x <- unlist(x)
    if (na.rm)
    x <- x[!is.na(x)]
    u <- unique(x)
    n <- length(u)
#get frequencies of each of the unique values in the vector
    frequencies <- rep(0, n)
    for (i in seq_len(n)) {
        if (is.na(u[i])) {
            frequencies[i] <- sum(is.na(x))
        }
        else {
            frequencies[i] <- sum(x == u[i], na.rm = TRUE)
        }
    }
#mode if a unimodal vector, else NA
    if (method == "mode" | is.na(method) | method == "")
        {return(ifelse(length(frequencies[frequencies==max(frequencies)])>1,NA,u[which.max(frequencies)]))}
#number of modes
    if (method == "modes" | method == "modevalues")
        {return(length(frequencies[frequencies==max(frequencies)]))}
#list of all modes
    if(method == "nmode" | method == "nmodes")
        {return(u[which(frequencies==max(frequencies), arr.ind = FALSE, useNames = FALSE)])}
#error trap the method
    warning("Warning: method not recognised.  Valid methods are 'mode' [default], 'nmodes' and 'modes'")
    return()
} #found at 'http://stackoverflow.com/revisions/15620958/1'
modeav(metadata$Weight, method="nmode")
```
The modes for weight were *__`r modeav(metadata$Weight, method="nmode")`__*

4.	Usign R commands, write the table to a new text file, but exclude the `Ear` and `Repro` columns.
```{r} 
metadata.ear <- metadata[,-9] #removes Ear column
metadata.repro <- metadata.ear[,-7] #removes Repro from metadata.ear
write.table(metadata.repro, file= "metadata_sansEarRepro.txt")
metadata.sansRE <- read.table(file="~/assignment02/metadata_sansEarRepro.txt", header=T)
metadata.sansRE[1:5, ]
```
