---
title: "assignment2"
output: html_document
---



This is Kaitlin's R Markdown file for Assignment 2

```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]
```

Assignment answers:

1. Data stats

- There are `r nrow(metadata)` samples described in this dataset.

- There are `r ncol(metadata)` columns in this dataset and their names are `r colnames(metadata)`.

```{r}
weight <- metadata[8]
weight15 <- weight[weight >= 15]
```

- There are `r length(weight15)` samples from mice that weighed 15 or more grams. 

- The median weight of the samples is `r apply(weight, 2, median)`. 

```{r}
sp4 <- metadata[4]
sp4 <- sp4[sp4 == "PMG"]
```

-There are `r length(sp4)` PMG samples. 

```{r}
#subsets metadata into two columns of interest
sp.sex <- subset(metadata, select = c("SP", "Sex"))
#indexes for conditionals
sselect <- subset(sp.sex, SP == "PL" & Sex == "F")
#creates number of selected outputs as a list to do length() on
s.select <- sselect[sselect == "PL"]
```

- There are `r length(s.select)` female PL samples.

```{r}
earorder <- order(metadata[[9]])
eardata <- metadata[earorder,]

```

- The matrix has been ordered by ear tag number and a sample is shown here 

```{r}
head(eardata)
```

```{r}
weightorder <- order(metadata[[8]])
weightdata <- metadata[weightorder,]

```

- The matrix has been ordered by weight and a sample is shown here:
```{r}
head(weightdata)
```

```{r}
stationorder <- order(metadata[[3]])
stationdata <- metadata[stationorder,]

temp <- table(as.vector(metadata[3]))
Nstation <- names(temp) [temp == max(temp)]
nmice <- subset(metadata, Station == Nstation)

```

- The most mice were captured at `r Nstation` and there were `r nrow(nmice)` captured there. 

The following code chunks perform the following operations:

This creates a sequence of numbers starting at 1, counting by 3s and going up to 100. 
```{r}
seq(1,100, 3)
```

Repeats a list of "a""b" ten times.
```{r}
rep(c("a","b"),10)
```

This generates 10 random deviations, order tells you which order they should be in 
```{r}
r <- runif(10); order(r)
```

Gives the remainder for 100 divided by 3
```{r}
100 %% 3
```

This code creates a matrix of the samples that have both weights of 16 and are SP PMG. 
```{r}
metadata[metadata$Weight==16 & metadata$SP=="PMG",]
```

Calculate the mode of weights for the metadata file:

```{r}
weighttemp <- table(as.vector(metadata[8]))
Nweight <- names(weighttemp) [weighttemp == max(weighttemp)]
```
 The mode is `r Nweight`. 

To write this table to a new file and exclude columns, I defined the columns we didn't want anymore as NULL

```{r}
metadata2 <- metadata 
metadata2$Repro <- NULL
metadata2$Ear <- NULL
write.table(metadata2, file = "noears.txt")
``` 





