---
title: "Assignment 2"
author: "Patrick D. Schloss"
date: "September 15, 2014"
output:
  html_document:
    keep_md: yes
---

Complete the exercises listed below and submit as a pull request to the [Assignment 2 repository](http://www.github.com/microbialinformatics/assignment02).  Format this document approapriately using R markdown and knitr. I would like to see your code blocks and output in the final documents you submit. As much as possible, you should output your solutions by embedding the solution within the text [see this example](https://github.com/microbialinformatics/assignment02/blob/master/example.Rmd). For those cases where there are multiple outputs, make it clear in how you format the text and interweave the solution, what the solution is.

Your pull request needs to include your *.Rmd and *.md files. Do not alter the `.gitignore` file. You may work with a partner, but you must submit your own assignment and give credit to anyone that worked with you on the assignment.

```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]
```

1.  Calculate the following on the data we read in from the `wild.metadata.txt` file that we discussed in class on 9/16/2014.

  * How many samples were described in the `wild.metadata.txt`?

  ```{r}
  nrow(metadata)
  ```
  
  There are `r nrow(metadata)` samples described in the `wild.metadata.txt`
  
  
  * How many columns are in the table? What are their names?
  ```{r}
  ncol(metadata)
  colnames(metadata)
  ```
  
  There are `r ncol(metadata)` columns in the table. The column names are `r colnames(metadata)`.
  
  
  * How many mice weighed 15 or more grams?
  ```{r}
  FifteenPlus<-subset(metadata,Weight>=15)
  FifteenPlus
  nrow(FifteenPlus)

  ```
  
  `r nrow(FifteenPlus)` mice weighed 15 or more grams
  
  * What is the median weight for the mice sampled?
  ```{r}
  median(metadata$Weight)
  ```
  
  The median weight of the mice sampled is `r median(metadata$Weight)`
  
  * How many PMG mice were there?
  ```{r}
  nrow(metadata[metadata$SP=="PMG",])
  ```
  
  There were `r nrow(metadata[metadata$SP=="PMG",])` *Peromyscus maniculatis* mice sampled. 
  
  * How many female PL mice were there? 
  ```{r}
  nrow(metadata[metadata$SP=="PL"&metadata$Sex=="F",])
  ```
  
  There were `r nrow(metadata[metadata$SP=="PL"&metadata$Sex=="F",])` female *Peromyscus leucopis* mice sampled.
  
  * Alphabetize `wild.metadata.txt` by the ear tag number (only show the first 5 rows of the table)
  ```{r}
 orderET<-order(metadata$ET)
 orderET5<-metadata[orderET,]
 head(orderET5,n=5)
  ```
 
  * Sort the table by the weight of each animal
  ```{r}
  metadata[order(metadata$Weight),]
  ```
  
  * The `Station` column indicates where the mice were sampled. Where were the most mice captured?
  
  ```{r}
  St.metadata<-metadata$Station
  st.metadata<-sort(table(St.metadata))
  ```
  Station N20 had the most mice collected.
  
  
  * How many mice were captured there?
  There were `r max(st.metadata)` mice collected at station N20.


2.	Describe what each of the following commands does in a text block above the code, be specific. Put the code into the appropriate knitr code chunk. If something throws an error or looks weird, fix it.

Output integers 1 through 100 in steps of 3.
```{r}
seq(1,100,3)
```

Repeat the character string "a" "b" 10 times.
```{r}
rep(c("a","b"),10)
```

Create a vector with the numbers 1 through 10 and order them randomly.
```{r}
r <- runif(10); order(r)
```

is "3" in "100"?
```{r}
100%in%3
```

Output any rows that contain a weight of 16 grams and are PMG species.
```{r}
metadata[metadata$Weight==16 && metadata$SP=="PMG",]
```


3.	Calculate the mode for the weight of the mice in `wild.metadata.txt`

```{r}
w.metadata<-metadata$Weight
Wmetadata<-(sort(table(w.metadata)))
MWmetadata<-tail(Wmetadata,n=2)
MWmetadata
```
There are two modes for the Weights

4.	Usign R commands, write the table to a new text file, but exclude the `Ear` and `Repro` columns.

```{r}
metadataShort<-metadata[,-9]
metadata.narrow<-metadataShort[,-7]
metadata.narrow
```
