---
title: "Assignment 2"
author: "Kat Wiles"
date: "September 26, 2014"
output:
  html_document:
    keep_md: yes
---

Data Source:
```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]
```

1.  Calculate the following on the data we read in from the `wild.metadata.txt` file that we discussed in class on 9/16/2014.

  * How many samples were described in the `wild.metadata.txt`?
  
```{r}
num.samples<-nrow(metadata)
```
  There were `r num.samples` samples
  
  
   * How many columns are in the table? What are their names
  
```{r}
num.columns<-ncol(metadata)
col.names<-colnames(metadata)
```  
  There are `r num.columns` columns with the following names: `r col.names`
  
  
  * How many mice weighed 15 or more grams?
```{r}
mice.weight15<-nrow(metadata[metadata$Weight>=15,])
```
    There were `r mice.weight15` mice that weighed 15 or more grams.
  
  * What is the median weight for the mice sampled?
```{r}
med.weight<-median(metadata$Weight)
```
      The median mice weight was `r med.weight` grams.
      
  * How many PMG mice were there?
  
```{r}
pmg.mice<-nrow(metadata[metadata$SP=="PMG",])
```
      There were `r pmg.mice` PMG mice sampled.
      
      
        * How many female PL mice were there?
```{r}
pl.mice.f<-nrow(metadata[metadata$SP=="PL" & metadata$Sex=="F",])
```  
      There were `r pl.mice.f` female PL mice.
      
      
 * Alphabetize `wild.metadata.txt` by the ear tag number (only show the first 5 rows of the table)
```{r}
metadata$ET<-factor(metadata$ET)
et.sort<-metadata[order(metadata$ET),]
et.sort[1:5,]
```
  
  
  
  * Sort the table by the weight of each animal
```{r}
metadata$Weight<-factor(metadata$Weight)
weight.sort<-metadata[order(metadata$Weight),]
weight.sort[1:5,]
``` 
  
  * The `Station` column indicates where the mice were sampled. Where were the most mice captured?
```{r}

best.station<-sort(table(metadata["Station"]), decreasing=TRUE)[1]
```
      The most mice were sampled at station`r names(best.station)`.
  
   
  * How many mice were captured there?
         
```{r}
N20mice<-metadata[metadata["Station"]=="N20",]
N20.unique<-N20mice[!duplicated(N20mice["ET"]),]
```
There were `r nrow(N20mice)` captures of `r nrow(N20.unique)` unique individuals at  `r names(best.station)`.


2.	Describe what each of the following commands does in a text block above the code, be specific. Put the code into the appropriate knitr code chunk. If something throws an error or looks weird, fix it.
Create a sequence from 1-100 by showing every 3 numbers.
```{r}
seq(1,100,3)
```


Repeat the sequence a b ten times.
```{r}
rep(c("a","b"),10)
```

Give a random distribution of 10 numbers and sort by their distribution.
```{r}
r <- runif(10); order(r)
```

What is 100 divided by 3?
```{r}
100 %/% 3
```

List all the PMG mice that weigh 16 grams?
```{r}
metadata[metadata$Weight==16 & metadata$SP=="PMG",]
```

3.	Calculate the mode for the weight of the mice in `wild.metadata.txt`


4.	Usign R commands, write the table to a new text file, but exclude the `Ear` and `Repro` columns.

