---
title: "Assignment 2"
author: "Patrick D. Schloss"
date: "September 15, 2014"
output:
  html_document:
    keep_md: yes
---

Complete the exercises listed below and submit as a pull request to the [Assignment 2 repository](http://www.github.com/microbialinformatics/assignment02).  Format this document approapriately using R markdown and knitr. I would like to see your code blocks and output in the final documents you submit. As much as possible, you should output your solutions by embedding the solution within the text [see this example](https://github.com/microbialinformatics/assignment02/blob/master/example.Rmd). For those cases where there are multiple outputs, make it clear in how you format the text and interweave the solution, what the solution is.

Your pull request needs to include your *.Rmd and *.md files. Do not alter the `.gitignore` file. You may work with a partner, but you must submit your own assignment and give credit to anyone that worked with you on the assignment.

```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]
```

1.  Calculate the following on the data we read in from the `wild.metadata.txt` file that we discussed in class on 9/16/2014.

  * How many samples were described in the `wild.metadata.txt`?   
  Answer:   
  ```{r}
n.samples <- nrow(metadata)
```  
     
There were `r n.samples` samples described in `wild.metadata.txt`.
  
  * How many columns are in the table? What are their names?     
  Answer:    
  ```{r}
n.col <- ncol(metadata)
names.col <- colnames(metadata)

```
     
There are `r n.col` columns in the table. And their names are: `r names.col`.
    
  * How many samples came from mice that weighed 15 or more grams?     
  Answer:   
  ```{r}
n.weight_15 <- nrow(metadata[metadata$Weight>=15,])

```
     
There are `r n.weight_15` mice weighed 15 or more grams.
  
  * What is the median weight for the mice sampled?     
  Answer:    
  ```{r}
median.weight <- median(metadata$Weight)

```
    
The median weight for the mice sampled is `r median.weight`.
  
  * How many PMG mice were there?    
  Answer:    
  ```{r}
n.PMG <- nrow(metadata[metadata$SP=="PMG",])

```
   
There are `r n.PMG` PMG mice.
  
  
  * How many female PL mice were there?      
  Answer:    
  ```{r}
n.f_PL <- nrow(metadata[metadata$SP=="PL" & metadata$Sex=="F",])

``` 
     
There are `r n.f_PL` female PL mice.
  
  
  * Alphabetize `wild.metadata.txt` by the ear tag number (only show the first 5 rows of the table)      
  Answer:    
  ```{r}
rows_ET <- head(metadata[order(metadata$ET),],5)

```
     
The first 5 rows of the table alphabetized by ear tag number are:   
    ```{r echo=FALSE} 
  rows_ET
 ```

  
  * Sort the table by the weight of the mice that each sample came from    
Answer:    
  
```{r}
o <- order(metadata$Weight)
z <- metadata
sort_by_weight <- z[o,]


```
     
  The table sorted by the weight of each animal is:   
    ```{r echo=FALSE} 
  sort_by_weight
 ```


  * The `Station` column indicates where the mice were sampled. Where were the most mice captured?    
Answer: 
  
```{r}
temp <- table(metadata$Station)
station_most <- names(temp)[temp == max(temp)]

```
     
 The station `r station_most` captured most mice.

  * How many mice were captured there?     
Answer:     
```{r}
n.mice_captured <- nrow(metadata[metadata$Station == "N20",])

```
    
`r n.mice_captured` mice were captured there.



2.	Describe what each of the following commands does in a text block above the code, be specific. Put the code into the appropriate knitr code chunk. If something throws an error or looks weird, fix it.

     
Output a sequence of integers from 1 to 100, and in jumps of 3.     
The result is:  `r seq(1,100,3)`

```
seq(1,100,3)
```
     

Replication. Replicate characters “a” and “b” for 10 times.    
The result is: `r rep(c("a","b"),10)`
```
rep(c("a","b"),10)
```
     


```
r <- runif(10); order(r)
```
      


```
100 % 3
```
      
      
Find out data in the table when weight equals 16, and SP is PMG at the same time.   
The result is:
```{r, echo=FALSE}
metadata[metadata$weight==16 && metadata$SP=="PMG",]
```
metadata[metadata$weight==16 && metadata$SP=="PMG",]

     

3.	Calculate the mode for the weight of the mice in `wild.metadata.txt`    
Answer:    
```{r}
table_w <- table(metadata$Weight)
mode_w <- names(table_w)[table_w == max(table_w)]

```
      
The mode for the weight of the mice in `wild.metadata.txt` is `r mode_w`.


4.	Usign R commands, write the table to a new text file, but exclude the `Ear` and `Repro` columns.   
Answer:    
```{r}
new.metadata <- metadata
new.metadata$Ear <- NULL
new.metadata$Repro <- NULL

```

