---

title: "Assignment 2"
author: "Shawn Whitefield"
date: "September 16, 2014"
output:
  html_document:
    keep_md: yes
  
---

Complete the exercises listed below and submit as a pull request to the [Assignment 2 repository](http://www.github.com/microbialinformatics/assignment02).  Format this document approapriately using R markdown and knitr. I would like to see your code blocks and output in the final documents you submit. Your pull request needs to include your *.Rmd and *.md files. Do not alter the `.gitignore` file. You may work with a partner, but you must submit your own assignment and give credit to anyone that worked with you on the assignment.

```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]
```

1.  Calculate the following on the data we read in from the `wild.metadata.txt` file that we discussed in class on 9/16/2014.

  * How many samples were described in the `wild.metadata.txt`?
  Answer: 111
```{r}
## getting summary information
summary(metadata)
# getting number of observations in metadata file
nrow(metadata)

```
 * How many columns are in the table? What are their names?

```{r}
#getting column number
ncol(metadata)
#getting column names
colnames(metadata)
```
* How many mice weighed 15 or more grams?
Answer: There are 77 observations from 31 mice with weights greater than or equal to 15 grams
```{r}
# There are repeat measurements for the same mouse
#First, assess how many individual mice are in the file
  # making ET a factor
factor(metadata$ET)
metadata$ET<-factor(metadata$ET)
#checking work
levels(metadata$ET)
summary(metadata)
head(metadata)
#There are 49 individual mice
#subsetting for weight >= 15, calling this set weightet
weightet<- subset(metadata, Weight>=15)
weightet
#check work
levels(weightet$ET)
#determine number of individual mice (ET) in weightet
uniquebigmice<- sort(weightet$ET)
unique(uniquebigmice)
     ```
```
What is the median weight for the mice sampled?
Answer: median weight among all observations is 16g
```{r}
#assuming this means meadian over all weight observations
median(metadata$Weight)
```
* How many PMG mice were there?
Answer: There are 53 observations of PMG from 23 mice
```{r}
#using summary feature to look at number in SP categories
summary(metadata)
#subsetting for only PMG values of SP
nPMG<- metadata[ which(metadata$SP=='PMG'), ]
#listing number of unique ET (mice) with PMG
sortnPMG<- sort(nPMG$ET)
sortnPMG
unique(sortnPMG)
```
* How many female PL mice were there?
Answer: There are 24 female PL mice
```{r}
nPL<- metadata[ which(metadata$SP=='PL'), ]
nPL
summary(nPL)
```
* Alphabetize `wild.metadata.txt` by the ear tag number (only show the first 5 rows of the table) 
```{r}

wild.metadata <- read.table(file="wild.metadata.txt", header=T)
sortbyet<-wild.metadata[order(metadata$ET),]
sortbyet[1:5,]

```

* Sort the table by the weight of each animal
```{r}
#sorting table by weight
byweighttable<-metadata[order(metadata$Weight),]
byweighttable
#there are multiple weights per animal(EarTag)
#sorting by ET and then by weight
byweight<-metadata[order(metadata$ET,metadata$Weight),]
byweight

```
* The `Station` column indicates where the mice were sampled. Where were the most mice captured? 
* How many mice were captured there?
Answer: N20, 4 mice.
```{r}
summary(metadata$Station)
```
2.  Describe what each of the following commands does in a text block above the code, be specific. Put the code into the appropriate knitr code chunk. If something throws an error or looks weird, fix it.
```{r}

```
count from 1 to 100 by 3's.
```
seq(1,100,3)
```
repeat vector containing a b, 10 times.
```
rep(c("a","b"),10)
```
give 10 numbers from 0 to 1 (from the uniform dist.) and put them in order
```
r <- runif(10); order(r)
```
Maybe this should be %*% instead of %? This is for matrix multiplication, so multiplying matrix with 100 by matrix with 3. 
```
100 %*% 3
```
Weight needs to have a capital W for the variable name. 
metadata[metadata$Weight==12 & metadata$SP=="PMG",]
```
metadata[metadata$weight==16 && metadata$SP=="PMG",]
```


3.	Calculate the mode for the weight of the mice in `wild.metadata.txt`
Answer: the weight is bimodal with 16 and 17 both occuring 12 times.
```{r}
#make variable containing a table with the weights as the lables and number of observations for each label in a vector.
weightfreq<- table(as.vector(metadata$Weight))
#check work
weightfreq
#ask for the names with the maximum frequency in the weightfreq vector.
names (weightfreq) [weightfreq == max(weightfreq)]
```

4.	Usign R commands, write the table to a new text file, but exclude the `Ear` and `Repro` columns.
```{r}
#make subset table without Ear and Repro
smallermetadata<- subset(metadata,select= c(-Ear, - Repro))
#check work
smallermetadata
#write table to text file
write.table(smallermetadata,"smallermetadata.txt")
```

