---
title: "Assignment 2"
author: "Isaiah Song"
date: "September 26, 2014"
output:
  html_document:
    keep_md: yes
---

Complete the exercises listed below and submit as a pull request to the [Assignment 2 repository](http://www.github.com/microbialinformatics/assignment02).  Format this document approapriately using R markdown and knitr. I would like to see your code blocks and output in the final documents you submit. As much as possible, you should output your solutions by embedding the solution within the text [see this example](https://github.com/microbialinformatics/assignment02/blob/master/example.Rmd). For those cases where there are multiple outputs, make it clear in how you format the text and interweave the solution, what the solution is.

Your pull request needs to include your *.Rmd and *.md files. Do not alter the `.gitignore` file. You may work with a partner, but you must submit your own assignment and give credit to anyone that worked with you on the assignment.

```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]
```

1.  Calculate the following on the data we read in from the `wild.metadata.txt` file that we discussed in class on 9/16/2014.

  * How many samples were described in the `wild.metadata.txt`?
  
```{r}
numSamples<-nrow(metadata) #Number of rows
```

There are as many samples as there are rows, equaling `r numSamples`.
  
  * How many columns are in the table? What are their names?
  
```{r}
numColumns<-ncol(metadata) #Number of columns
namColumns<-colnames(metadata) #Names of columns

```

There are `r numColumns` columns and their names are: `r namColumns`.

  * How many samples came from mice that weighed 15 or more grams?
  
```{r}
wtGr15<-metadata[metadata$Weight>=15,] #Output of samples weighing 15 grams or greater
samGr15<-nrow(wtGr15) #Number of rows
```

The number of samples that came from mice that weighed 15 or more grams is `r samGr15`.
  
  * What is the median weight of the samples?
  
```{r}
medWt<-median(metadata$Weight) #Median of sample weights
```

The median weight is `r medWt`.

  * How many PMG samples were there?
  
```{r}
PMG<-metadata[metadata$SP=="PMG",] #Output of samples of PMG
numPMG<-nrow(PMG) #Number of rows
```

There were `r numPMG` PMG samples.

  * How many female PL samples were there?
  
```{r}
PL_F<-metadata[metadata$SP=="PL"&metadata$Sex=="F",] #Output of samples of female PL
numPL_F<-nrow(PL_F) #Number of rows
```

There were `r numPL_F` female PL samples.

  * Alphabetize `wild.metadata.txt` by the ear tag number (only show the first 5 rows of the table) 
  
```{r}
metadata.sortET<-metadata[order(metadata$ET),] #Sort by ET order
metadata.sortET[1:5,] #Display first 5 rows
```

  * Sort the table by the weight of the mice that each sample came from  
  
```{r}
metadata.sortWt<-metadata[order(metadata$Weight),] #Sort by weight
metadata.sortWt[1:5,] #Display first 5 rows
```

  * The `Station` column indicates where the mice were sampled. Where were the most mice captured? 
  
```{r}
summStation<-summary(metadata$Station) #Show number of mice sampled in each station
sort(summStation) #Sort by number of mice sampled
```

The most mice were captured in N20.

  * How many mice were captured there?

```{r}
sortStation<-sort(summStation, decreasing=TRUE) #Sort by decreasing number of mice sampled
sortStation[1] #Output first entry (station of highest number of mice sampled)
```

There were 4 mice captured at N20.

2.	Describe what each of the following commands does in a text block above the code, be specific. Put the code into the appropriate knitr code chunk. If something throws an error or looks weird, fix it.


Gives a sequence of numbers from 1 to 100 rising in increments of 3.
```
seq(1,100,3)
```

Repeats a vector sequence ("a" and "b") a total of 10 times
```
rep(c("a","b"),10)
```

Gives a series of 10 numbers in uniform distribution from 0 to 1 and sorts them in order
```
r <- runif(10); order(r)
```

This code gives an error; may possibly be fixed by changing to "100 %% 3", which gives the remainder when dividing the first number by the second number
```
100 % 3
```

Outputs rows in the metadata that correspond to a weight of 16 and SP value of "PMG";  Also, "weight" must be capitalized as that is how the variable is assigned
```
metadata[metadata$weight==16 && metadata$SP=="PMG",]
```


3.	Calculate the mode for the weights in `wild.metadata.txt`


4.	Usign R commands, write the table to a new text file, but exclude the `Ear` and `Repro` columns.

