---
title: "Assignment 2"
author: "Patrick D. Schloss"
date: "September 15, 2014"
output:
  html_document:
    keep_md: yes
---

Complete the exercises listed below and submit as a pull request to the [Assignment 2 repository](http://www.github.com/microbialinformatics/assignment02).  Format this document approapriately using R markdown and knitr. I would like to see your code blocks and output in the final documents you submit. As much as possible, you should output your solutions by embedding the solution within the text [see this example](https://github.com/microbialinformatics/assignment02/blob/master/example.Rmd). For those cases where there are multiple outputs, make it clear in how you format the text and interweave the solution, what the solution is.

Your pull request needs to include your *.Rmd and *.md files. Do not alter the `.gitignore` file. You may work with a partner, but you must submit your own assignment and give credit to anyone that worked with you on the assignment.

```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]
```

1.  Calculate the following on the data we read in from the `wild.metadata.txt` file that we discussed in class on 9/16/2014.

  * How many samples were described in the `wild.metadata.txt`?
```{r}
n.samples <- nrow(metadata[,])
```
<!---
Does this count include headers?
-->
There were `r n.samples` samples described in the `wild.metadata.txt`.
  
  * How many columns are in the table? What are their names? 
```{r}
n.columns <- ncol(metadata[,])
names <- colnames(metadata)
```
There are `r n.columns` columns in the table.
Their names are `r names`.

  * How many mice weighed 15 or more grams?
```{r}
more.weight <- metadata$Weight >= 15
n.more.weight <- nrow(metadata[more.weight,])
```
There were `r n.more.weight` mice that weighed 15 or more grams.

  * What is the median weight for the mice sampled?
```{r}
med <- median(metadata$Weight, na.rm = FALSE)
```
The median weight for the mice sampled was `r med`.

  * How many PMG mice were there?
```{r}
PMG <- metadata$SP == "PMG"
n.PMG <- nrow(metadata[PMG,])
```
There were `r n.PMG` PMG mice.

  * How many female PL mice were there?
```{r}
PL.F <- metadata$SP == "PL" & metadata$Sex == "F"
n.PL.F <- nrow(metadata[PL.F,])
```
There were `r n.PL.F` mice.

  * Alphabetize `wild.metadata.txt` by the ear tag number (only show the first 5 rows of the table)
```{r}
sorted.metadata <- metadata[order(metadata$ET), ]
```
The first 5 rows of the sorted table are:
```{r, echo=FALSE}
sorted.metadata[1:5,]
```

  * Sort the table by the weight of each animal
```{r}
sortw.metadata <- metadata[order(metadata$Weight), ]
```
The first 5 rows of the sorted table are:
```{r, echo=FALSE}
sortw.metadata[1:5,]
```

  * The `Station` column indicates where the mice were sampled. Where were the most mice captured?
  
  * How many mice were captured there?
```{r}
many.mice <- metadata$Station == "N20"
n.many.mice <- nrow(metadata[many.mice,])
```
The most mice were captured at the blank station.
There were `r n.many.mice` mice captured there.

2.	Describe what each of the following commands does in a text block above the code, be specific. Put the code into the appropriate knitr code chunk. If something throws an error or looks weird, fix it.

This command will generate a vector of values that is every third number from 1 to 100.
```{r}
seq(1,100,3)
```

```
rep(c("a","b"),10)
```

```
r <- runif(10); order(r)
```

```
100 % 3
```

```
metadata[metadata$weight==16 && metadata$SP=="PMG",]
```


3.	Calculate the mode for the weight of the mice in `wild.metadata.txt`


4.	Usign R commands, write the table to a new text file, but exclude the `Ear` and `Repro` columns.

