---
title: "Assignment 2"
author: "Patrick D. Schloss"
date: "September 15, 2014"
output:
  html_document:
    keep_md: yes
---

Complete the exercises listed below and submit as a pull request to the [Assignment 2 repository](http://www.github.com/microbialinformatics/assignment02).  Format this document approapriately using R markdown and knitr. I would like to see your code blocks and output in the final documents you submit. As much as possible, you should output your solutions by embedding the solution within the text [see this example](https://github.com/microbialinformatics/assignment02/blob/master/example.Rmd). For those cases where there are multiple outputs, make it clear in how you format the text and interweave the solution, what the solution is.

Your pull request needs to include your *.Rmd and *.md files. Do not alter the `.gitignore` file. You may work with a partner, but you must submit your own assignment and give credit to anyone that worked with you on the assignment.

```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
rownames(metadata) <- metadata$Group
metadata <- metadata[,-1]
```

1.  Calculate the following on the data we read in from the `wild.metadata.txt` file that we discussed in class on 9/16/2014.

  * How many samples were described in the `wild.metadata.txt`?
  
```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
n.row <- nrow(metadata) 
```
There were `r n.row` samples described in the `wild.metadata.txt`
  
  * How many columns are in the table? What are their names?
```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
n.col <- ncol(metadata)
name.col <- colnames(metadata)
```
There were `r n.col` columns in the table, and their names are `r name.col` 

  * How many mice weighed 15 or more grams?  
```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
w <- metadata$Weight
w15 <- w[w>15]
n.w15 <- length(w15)

```
There were `r n.w15` mice weighed 15 or more grams.

  * What is the median weight for the mice sampled?  
```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
w <- metadata$Weight
median.W <- median(w)
```
The median weight for the mice sampled is `r median.W`. 
 
 * How many PMG mice were there?
```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
pmg <- metadata[metadata$SP=="PMG",]
n.pmg <- nrow(pmg)
``` 
 There were `r n.pmg` mice.
 
 * How many female PL mice were there?
```{r}
metadata <- read.table(file="wild.metadata.txt", header=T)
f.pl <- metadata[metadata$SP=="PL"& metadata$Sex=="F",]
n.fpl <- nrow(f.pl)
```
There were `r n.fpl` mice.
  
  * Alphabetize `wild.metadata.txt` by the ear tag number (only show the first 5 rows of the table)
```{r}
o.et <- order(metadata$ET)
oet <- metadata[o.et,]
head(oet)
```
The first 5 rows of the table that is alphabetized by ear tag number.
  
  * Sort the table by the weight of each animal
```{r}
metadata[order(metadata$Weight),]
```
  
  * The `Station` column indicates where the mice were sampled. Where were the most mice captured?
```{r}
a <- metadata$Station
sort(table(a), decreasing = TRUE)[1]

```
The station where the most mice were captured is `r names(sort(table(a), decreasing = TRUE)[1])` .
  
  * How many mice were captured there?

And `r sort(table(a), decreasing = TRUE)[1]` mice were captured there.

2.	Describe what each of the following commands does in a text block above the code, be specific. Put the code into the appropriate knitr code chunk. If something throws an error or looks weird, fix it.

>create a sequence of equidistant numbers in jumps of 3
```{r}
seq(1,100,3)
```

>generate 10 replicates of vectors that contains string variables a and b 
```{r}
rep(c("a","b"),10)
```

>generate a vector of 10 random numbers and get the elements in the vector in the correct order
```{r}
r <- runif(10); order(r)
```

>100 mod 3
```{r}
100 %% 3
```

>get the data for samples that Weight 16 grams and have PMG.
```{r}
metadata[metadata$Weight==16 & metadata$SP=="PMG",]
```


3.	Calculate the mode for the weight of the mice in `wild.metadata.txt`
```{r}
w <- metadata$Weight
temp <- table(w)
names(temp)[temp == max(temp)]
```
4.	Usign R commands, write the table to a new text file, but exclude the `Ear` and `Repro` columns.
```{r}
metadata$Ear <- NULL
metadata$Repro <- NULL
new.metadata <- metadata
write.table(new.metadata, "c:/Users/Yao/Documents/assignment02/new.metadata.txt", sep="\t")
```
